---
title: "Tutorial 1: Handling social network data in R"
author: "Bjarne van der Mark"
output:
  html_document:
    number_sections: false
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(knitr)
library(igraph)
library(dplyr)
library(reshape2)
library(tidyverse)
library(network)

knitr::opts_chunk$set(
  echo = TRUE,
  tidy.opts = list(width.cutoff = 100),
  tidy = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  cache = TRUE,
  class.source = c("test"),
  class.output = c("test2")
)

options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {
  sprintf("<span style='color: %s;'>%s</span>", color, x)
}
```

```{r klippy, echo=FALSE, include=TRUE}
# klippy::klippy(position = c('top', 'right'))
# klippy::klippy(color = 'darkred')
# klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

On this page you will find my code from tutorial 1.

## 1.1
### 1.1.1

```{r}
url1 <- "https://github.com/rensec/sasr08/raw/main/g_adj_matrix_simple.csv"
g_matrix <- read.csv(file = url1, header = FALSE)
g_matrix

class(g_matrix)

g_matrix <- as.matrix(g_matrix)
g_matrix

rownames(g_matrix) <- 1:nrow(g_matrix)
colnames(g_matrix) <- 1:ncol(g_matrix)
g_matrix
```

Question: How many nodes are included in this matrix, and how many edges are there between these nodes?

Answer: There are five nodes in this matrix and four edges.

```{r}
g_nodes_age <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_nodes_age.csv")
g_nodes_age

g <- graph_from_adjacency_matrix(g_matrix)
class(g)
g

plot(g)

g <- set_vertex_attr(
  g,
  name = "age",
  value = g_nodes_age$age
)
```

### 1.1.2

```{r}
g_elist <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_elist.csv")
g_elist

g_from_elist <- graph_from_data_frame(g_elist)
plot(g_from_elist)

g_from_elist <- graph_from_data_frame(g_elist, vertices = g_nodes_age)
plot(g_from_elist)
```

### Assignment: Edge list pipeline

From the code snippets above, construct a complete pipeline to go from the edge list to the igraph object. Start from `g_elist`. Plot the resulting network. Store the code somewhere for later use.

```{r}
g_from_elist <- graph_from_data_frame(
  d = g_elist,
  vertices = g_nodes_age
)

plot(g_from_elist)
```

### 1.1.3

```{r}
g_adj_list <- read.csv(file = "https://github.com/rensec/sasr08/raw/main/g_adj_list.csv")
g_adj_list

g_elist_from_alist <- g_adj_list

g_elist_from_alist <- g_adj_list %>%
  select(id, friend1:friend2) %>%
  melt(id.vars = "id")

g_adj_list %>%
  select(id, friend1:friend2) %>%
  pivot_longer(c(friend1, friend2))

g_elist_from_alist <- g_elist_from_alist %>%
  filter(!is.na(value)) %>%
  rename(from = "id", to = "value", sourcevar = "variable") %>%
  relocate(to, .after = from)

g_elist_from_alist
```

Question: Besides being useful for data manipulation procedures, keeping the information in `sourcevar` could also be important for more substantive reasons. Can you think of such a reason?

Answer: Seeing who is a more important friend.

```{r}
g_from_alist <- graph_from_data_frame(g_elist_from_alist)
plot(g_from_alist)

nodelist <- select(g_adj_list, id, age)

g_from_alist <- graph_from_data_frame(
  g_elist_from_alist,
  vertices = nodelist
)

plot(g_from_alist)
```

### Assignment: Adjacency list pipeline

From the code snippets above, construct a complete pipeline to go from the adjacency list to the igraph object. Start from `g_adj_list`. Plot the resulting network. Store the code for later use.

```{r}
g_from_alist <- graph_from_data_frame(
  d = g_adj_list %>%
    select(id, friend1:friend2) %>%
    pivot_longer(c(friend1, friend2)) %>%
    filter(!is.na(value)) %>%
    rename(from = id, to = value, sourcevar = name) %>%
    relocate(to, .after = from),
  
  vertices = g_adj_list %>%
    select(id, age),
  
  directed = TRUE
)

plot(g_from_alist)
```

## 1.2
### 1.2.1

```{r}
g_adj_matrix <- as_adjacency_matrix(g, sparse = FALSE)
g_adj_matrix

all(g_adj_matrix == g_matrix)
```

### 1.2.2

```{r}
g_edgelist <- igraph::as_data_frame(g_from_elist, what = "edges")
g_edgelist

igraph::as_data_frame(g_from_elist, what = "vertices")
```

### 1.2.3

```{r}
edgelist_2 <- igraph::as_data_frame(g_from_alist, what = "edges")
nodelist_2 <- igraph::as_data_frame(g_from_alist, what = "vertices")

d <- edgelist_2 %>%
  rename(id = "from") %>%
  pivot_wider(
    id_cols = id,
    names_from = sourcevar,
    values_from = to
  ) %>%
  merge(nodelist_2, by.x = "id", by.y = "name", all.y = TRUE) %>%
  relocate(age, .after = id) %>%
  type_convert()

d
```

## 1.3

```{r}
g_np <- as.network(
  g_elist_from_alist,
  matrix.type = "edgelist",
  vertices = nodelist
)

g_np

plot(g_np)
```

## 1.4

```{r}
g_mod <- delete_vertices(g, 2)
plot(g_mod)

g_mod <- delete_vertices(g, which(V(g)$age == 21))
plot(g_mod)
```

### Question

Write the code to remove all nodes from `g` for which age is missing.

```{r}
g_mod <- delete_vertices(g, which(is.na(V(g)$age)))
```

```{r}
V(g)$gender <- c("male", "female", "female", "other", "male")
g

g_rev <- reverse_edges(g)
plot(g_rev)

g_union <- igraph::union(g, g_rev)
```

Question: Before you run `plot(g_union)`, think about what the resulting network should look like. Thus, what could you use the combination of `reverse_edges()` and `union()` for?

Answer: You can use this to make the network undirected.

```{r}
g_undir <- as_undirected(g)
plot(g_undir)
```

Question: Realize however that, conceptually, `g_union` and `g_undir` are different. What is the difference?

Answer: `g_union` goes both directions and `g_undir` has no directions.